---
title: "Example using Python and R"
author: "Nick Malleson"
date: "018/07/2017"
output:
  html_document: 
    toc: yes
    pandoc_args: [
      "+RTS", "-K64m",
      "-RTS"
    ]
  pdf_document:
    fig_crop: no
    highlight: kate
    keep_tex: yes
    latex_engine: xelatex
    number_sections: yes
fontsize: 10pt
---

A small example of how to use R and Python together in the same R Markdown document

This uses the [feather](https://blog.rstudio.com/2016/03/29/feather/) library that allows R and python to share data by writing binary files.

## Initialise

```{r init}
library(feather) # To allow python and R to share data
```

## Create some data with R

For this example we'll just create a list of random numbers to be manipulated in R and Python (code frome [here](https://stackoverflow.com/questions/19352267/r-create-a-data-frame-with-random-numbers-in-each-column)).

Make a data frame with two columns and 1000 rows.

```{r createData }

data <- data.frame(replicate(2,sample(0:10,1000,rep=TRUE)))

par(mfrow=c(2,1))
hist(data$X1)
hist(data$X2)

```

## Write the data using feather

Write out the `data` dataframe using Feather so that it can be read in and manipulated by Python.

```{r writeToFeatherFromR }

write_feather(data, path="./data.feather")

```

## Read and process the data in python

Now the neat bit - use python to read the data from R, do something with it (we'll just make a new column) and then write it back out.

Note that in the R Markdown document, you can specify a particular python engine to use. E.g. the following would use an anaconda environment called 'py35':

```
{python engine.path="/Users/nick/anaconda2/envs/py35/bin/python"}
```

You'll have to change that to match your own setup if you want to run this script.

Which ever python environment you use, you have to have the `feather` library installed. You can install it as normal, with something like: 

```pip install feather-format```

```{python readInPython, engine.path="/Users/nick/anaconda2/envs/py35/bin/python" }

import feather

# Read in the data frame
data = feather.read_dataframe("./data.feather")

# Create a new column by multiplying the first two columns:
data.loc[:,'X3'] = data.loc[:,'X1'] * data.loc[:,'X2']

# Write out the data using feather
feather.write_dataframe(data, "./data.feather")

```

## Re-read the data in R

Now that we've processed it in python, re-read it using R.

```{r reReadData }

new.data <- read_feather("./data.feather")

hist(new.data$X3)

```

That's it!

OK that's a very simple example, but the ability to use either language in the same R Markdown document, and to share data, is extremely powerful.